---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dwelch

<!-- badges: start -->
<!-- badges: end -->

<tt>dwelch</tt> provides code to calculate the debiased Welch estimator developed 
[here](https://arxiv.org/pdf/2312.13643.pdf). Code for routine data manipulation 
and plotting is hidden; if you are interested in this consult <tt>README.Rmd</tt>. 
<tt>python</tt> implementation of the code is part of the larger <tt>speccy</tt> package 
available [here](https://github.com/TIDE-ITRH/speccy).


## Installation

You can install the development version of <tt>dwelch</tt> from [GitHub](https://github.com) with:

``` r
devtools::install_github("astfalckl/dwelch")
```

## Generate AR(4) Example

We demonstrate the basic functionality of <tt>dwelch</tt> with the classic AR(4) problem. First, import some basic packages.

```{r message = FALSE}
library(dwelch)
library(ggplot2)
library(tidyverse)
library(latex2exp)
library(patchwork)
```

```{r echo = FALSE}
set.seed(91)
theme_set(theme_bw())
```

Next, we set our parameters and generate an AR process.

```{r}
m <- 2^6
l <- 2^9
overlap <- 0
s <- l * (1 - overlap)
n <- (m - 1) * s + l
delta <- 1

phis <- c(2.7607, -3.8106, 2.6535, -0.9238)
sd <- 1

sampled_ar <- stats::arima.sim(
    list(ar = phis), n, n.start = 1000, sd = sd
)
```

Define our data taper, <tt>h</tt>, and calculate Welch's estimate of the AR process. We will show results for 
the boxcar and Hamming tapers side-by-side. See package [gsignal](https://cran.r-project.org/web/packages/gsignal/index.html)
for a fairly comprehensive list of other tapers. Note, gsignal masks the function <tt>pwelch</tt> and so if
gsignal is loaded into your R session you will need to explicitly call <tt>dwelch::pwelch</tt>.

```{r}
h_hm <- gsignal::hamming(l) # Hamming filter
h_bc <- rep(1, l) #Boxcar filter

pwelch_bc <- dwelch::pwelch(sampled_ar, m, l, s, delta, h_bc)
pwelch_hm <- dwelch::pwelch(sampled_ar, m, l, s, delta, h_hm)
```

```{r, echo = FALSE, fig.width = 10, fig.height = 5}
ff <- seq(1, l - 1) / l
ff <- ff[ff < 0.5]
proc_var <- 2 * sum(ar_spectrum(ff, phis, sd, delta) * diff(ff)[1])
ar_acf <- stats::ARMAacf(ar = phis, lag.max = l - 1)

biased_bc_plot <- tibble(
    ff = ff,
    true_ar_spec = ar_spectrum(ff, phis, sd, delta),
    biased_ar_spec = dwelch::bochner(proc_var * ar_acf, h = h_bc),
    pwelch = pwelch_bc$pwelch
    ) %>%
    ggplot() +
        geom_line(
            aes(x = ff, y = true_ar_spec),
            linetype = "dashed", alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = biased_ar_spec),
            alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = pwelch)
        ) +
        scale_y_continuous(
            TeX("Spectral density [$m^2 / Hz]"),
            trans = "log10",
            limits = c(0.7 * 10^-2, 2.5 * 10^4),
            breaks = c(10^-2, 10^0, 10^2, 10^4),
            labels = c(
                TeX("$10^{-2}$"),
                TeX("$10^0$"),
                TeX("$10^2$"),
                TeX("$10^4$")
            )) +
        scale_x_continuous("Frequency [Hz]") +
        ggtitle("Rectangular Taper") +
        theme_bw() +
        theme(
            panel.grid.minor = element_blank()
        )

biased_hm_plot <- tibble(
    ff = ff,
    true_ar_spec = ar_spectrum(ff, phis, sd, delta),
    biased_ar_spec = dwelch::bochner(proc_var * ar_acf, h = h_hm),
    pwelch = pwelch_hm$pwelch
    ) %>%
    ggplot() +
        geom_line(
            aes(x = ff, y = true_ar_spec),
            linetype = "dashed", alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = biased_ar_spec),
            alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = pwelch)
        ) +
        scale_y_continuous(
            TeX("Spectral density [$m^2 / Hz]"),
            trans = "log10",
            limits = c(0.7 * 10^-2, 2.5 * 10^4),
            breaks = c(10^-2, 10^0, 10^2, 10^4),
            labels = c(
                TeX("$10^{-2}$"),
                TeX("$10^0$"),
                TeX("$10^2$"),
                TeX("$10^4$")
            )) +
        scale_x_continuous("Frequency [Hz]") +
        ggtitle("Hamming Taper") +
        theme_bw() +
        theme(
            panel.grid.minor = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y = element_blank()
        )

pwelch_plot <- biased_bc_plot + biased_hm_plot
pwelch_plot
```

## Calculate debiased Welch estimate

For the debiased Welch estimator we must only make one additional specification: the number of debiased 
bases, <tt>k</tt>. Aside from the selection of <tt>k</tt>, the function <tt>dwelch</tt> excutes similarly 
to <tt>pwelch</tt>, above.

```{r}
dwelch_bc <- dwelch::dwelch(
    sampled_ar, m, l, s, k = l / 4, h = h_bc, model = "nnls"
)
dwelch_hm <- dwelch::dwelch(
    sampled_ar, m, l, s, k = l / 4, h = h_hm, model = "nnls"
)
```

```{r, echo = FALSE, fig.width = 10, fig.height = 5}
debiased_bc_plot <- tibble(
    ff = ff,
    true_ar_spec = ar_spectrum(ff, phis, sd, delta),
    biased_ar_spec = dwelch::bochner(ar_acf)
) %>%
    ggplot() +
        geom_line(
            aes(x = ff, y = true_ar_spec),
            linetype = "dashed", alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = proc_var * biased_ar_spec),
            alpha = 0.25, size = 0.75
        ) +
        geom_line(
            data = dwelch_bc %>%
                filter(dwelch != 0),
            aes(x = ff, y = dwelch)
        ) +
        scale_y_continuous(
            TeX("Spectral density [$m^2 / Hz]"),
            trans = "log10",
            limits = c(0.7 * 10^-2, 2.5 * 10^4),
            breaks = c(10^-2, 10^0, 10^2, 10^4),
            labels = c(
                TeX("$10^{-2}$"),
                TeX("$10^0$"),
                TeX("$10^2$"),
                TeX("$10^4$")
            )) +
        scale_x_continuous("Frequency [Hz]") +
        ggtitle("Rectangular Taper, Debiased") +
        theme_bw() +
        theme(
            panel.grid.minor = element_blank()
        )

debiased_hm_plot <- tibble(
    ff = ff,
    true_ar_spec = ar_spectrum(ff, phis, sd, delta),
    biased_ar_spec = dwelch::bochner(proc_var * ar_acf, h = h_hm)
) %>%
    ggplot() +
        geom_line(
            aes(x = ff, y = true_ar_spec),
            linetype = "dashed", alpha = 0.25, size = 0.75
        ) +
        geom_line(
            aes(x = ff, y = biased_ar_spec),
            alpha = 0.25, size = 0.75
        ) +
        geom_line(
            data = dwelch_hm %>%
                filter(dwelch != 0),
            aes(x = ff, y = dwelch)
        ) +
        scale_y_continuous("Spectral density", trans = "log10",
            limits = c(0.7 * 10^-2, 2.5 * 10^4),
            breaks = c(10^-2, 10^0, 10^2, 10^4),
            labels = c(
                TeX("$10^{-2}$"),
                TeX("$10^0$"),
                TeX("$10^2$"),
                TeX("$10^4$")
            )) +
        scale_x_continuous("Frequency [Hz]") +
        ggtitle("Hamming Taper, Debiased") +
        theme_bw() +
        theme(
            panel.grid.minor = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y = element_blank()
        )

dwelch_plot <- debiased_bc_plot + debiased_hm_plot
dwelch_plot
```

## Non-negative solutions

The WLS solution does not constain the debiased estimator to non-negative solutions, which is required of a 
spectral estimator. Note, this happens when the signal at a frequency is masked by spectral leakage.
<tt>dwelch</tt> has functionality to constrain solutions to be non-negative, this is done by setting
<tt>model = "nnls"</tt>. See the example below where we have selected an example with a particularly bad 
solution space. Note, non-positive values are not plotted. Note that for the Hamming tapered data where 
the solution is already non-negative, the nnls solution is the same.

```{r, fig.width = 10, fig.height = 5}
set.seed(23)

k = l / 4

sampled_ar <- stats::arima.sim(list(ar = phis), n, n.start = 1000, sd = sd)

pwelch_bc <- dwelch::pwelch(sampled_ar, m, l, s, delta, h_bc)
pwelch_hm <- dwelch::pwelch(sampled_ar, m, l, s, delta, h_hm)

dwelch_bc <- dwelch(sampled_ar, m, l, s, k, delta, h_bc)
dwelch_hm <- dwelch(sampled_ar, m, l, s, k, delta, h_hm)

nnls_bc <- dwelch(sampled_ar, m, l, s, k, delta, h_bc, model = "nnls")
nnls_hm <- dwelch(sampled_ar, m, l, s, k, delta, h_hm, model = "nnls")
```

```{r, echo = FALSE, warning = FALSE, fig.width = 10, fig.height = 5}
pwelch_both <- pwelch_bc %>%
    left_join(pwelch_hm, by = "ff", suffix = c("_Boxcar", "_Hamming")) %>%
    pivot_longer(
        -ff,
        names_to = "filter",
        names_prefix = "pwelch_",
        values_to = "pwelch"
    )

dwelch_both <- dwelch_bc %>%
    left_join(dwelch_hm, by = "ff", suffix = c("_Boxcar", "_Hamming")) %>%
    pivot_longer(
        -ff,
        names_to = "filter",
        names_prefix = "dwelch_",
        values_to = "dwelch"
    )

nnls_both <- nnls_bc %>%
    left_join(nnls_hm, by = "ff", suffix = c("_Boxcar", "_Hamming")) %>%
    pivot_longer(
        -ff,
        names_to = "filter",
        names_prefix = "dwelch_",
        values_to = "nnls"
    )

pwelch_both %>%
    mutate(ar = ar_spectrum(pwelch_both$ff, phis, sd, delta)) %>%
    ggplot() +
    geom_line(aes(x = ff, y = ar), linewidth = 1, alpha = 0.3) +
    geom_line(aes(x = ff, y = pwelch, colour = "pwelch")) +
    geom_line(
        data = nnls_both %>% filter(nnls > 1e-8),
        mapping = aes(x = ff, y = nnls, colour = "nnls")
    ) +
    geom_line(
        data = dwelch_both,
        mapping = aes(x = ff, y = dwelch, colour = "dwelch")
    ) +
    scale_x_continuous(
        "Frequency [Hz]",
        limits = c(0, 0.5 / delta),
        expand = c(0, 0)
    ) +
    scale_y_continuous(
        "Spectral density",
        limits = c(1e-3, 1e5),
        expand = c(0, 0),
        trans = "log10"
    ) +
    scale_color_manual(
        values = c(
            "pwelch" = "#002385",
            "dwelch" = "#8d3b00",
            "nnls" = "#00513b"
        )
    ) +
    facet_wrap(vars(filter)) +
    theme(
        legend.position = c(0.9, 0.8),
        legend.title = element_blank(),
        text = element_text(size = 16)
    )
```

## Convergence to Welch's estimate

```{r}
set.seed(28)

k <- round(get_nfreq(l), 0)

m1 <- 16
m2 <- 64
n1 <- (m1 - 1) * s + l
n2 <- (m2 - 1) * s + l

h_hn <- gsignal::hann(l) # Hann filter

sampled_ar1 <- stats::arima.sim(list(ar = phis), n1, n.start = 1000, sd = sd)
sampled_ar2 <- stats::arima.sim(list(ar = phis), n2, n.start = 1000, sd = sd)

pwelch_hn1 <- dwelch::pwelch(sampled_ar1, m1, l, s, delta, h_hn)
dwelch_hn1 <- dwelch::dwelch(sampled_ar1, m1, l, s, k, delta, h_hn)
pwelch_hn2 <- dwelch::pwelch(sampled_ar2, m2, l, s, delta, h_hn)
dwelch_hn2 <- dwelch::dwelch(sampled_ar2, m2, l, s, k, delta, h_hn)
```

```{r, echo = FALSE, warning = FALSE, fig.width = 10, fig.height = 5}

annotation_custom2 <- function(
    grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data
) {
    layer(
        data = data,
        stat = StatIdentity,
        position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE,
        params = list(
            grob = grob,
            xmin = xmin,
            xmax = xmax,
            ymin = ymin,ymax = ymax
        )
    )
}

pwelch_both <- pwelch_hn1 %>%
    left_join(pwelch_hn2, by = "ff", suffix = c("_m=16", "_m=64")) %>%
    pivot_longer(
        -ff,
        names_to = "m",
        names_prefix = "pwelch_",
        values_to = "pwelch"
    )

dwelch_both <- dwelch_hn1 %>%
    left_join(dwelch_hn2, by = "ff", suffix = c("_m=16", "_m=64")) %>%
    pivot_longer(
        -ff,
        names_to = "m",
        names_prefix = "dwelch_",
        values_to = "dwelch"
    )

p1 <- pwelch_both %>%
    mutate(ar = ar_spectrum(pwelch_both$ff, phis, sd, delta)) %>%
    ggplot() +
    geom_line(aes(x = ff, y = ar), linewidth = 1, alpha = 0.3) +
    geom_line(aes(x = ff, y = pwelch, colour = "pwelch")) +
    geom_line(
        data = dwelch_both,
        mapping = aes(x = ff, y = dwelch, colour = "dwelch")
    ) +
    scale_x_continuous(
        "Frequency [Hz]",
        limits = c(0, 0.5 / delta),
        expand = c(0, 0)
    ) +
    scale_y_continuous(
        "Spectral density",
        limits = c(1e-3, 1e5),
        expand = c(0, 0),
        trans = "log10"
    ) +
    scale_color_manual(
        values = c("pwelch" = "#002385", "dwelch" = "#8d3b00")
    ) +
    facet_wrap(vars(m)) +
    theme(legend.position = c(0.93, 0.9), legend.title = element_blank())

p2a <- pwelch_both %>%
    mutate(ar = ar_spectrum(pwelch_both$ff, phis, sd, delta)) %>%
    filter(m == "m=16") %>%
    ggplot() +
    geom_line(aes(x = ff, y = ar), linewidth = 1, alpha = 0.3) +
    geom_line(aes(x = ff, y = pwelch, colour = "pwelch")) +
    geom_line(
        data = dwelch_both %>% filter(m == "m=16"),
        mapping = aes(x = ff, y = dwelch, colour = "dwelch")
    ) +
    scale_x_continuous(
        "Frequency [Hz]",
        limits = c(0.25, 0.35),
        expand = c(0, 0)
    ) +
    scale_y_continuous(
        "Spectral density",
        limits = c(1e-2, 5e-1),
        expand = c(0, 0),
        trans = "log10"
    ) +
    scale_color_manual(
        values = c("pwelch" = "#002385", "dwelch" = "#8d3b00")
    ) +
    theme(legend.position = "none", axis.title = element_blank())

p2b <- pwelch_both %>%
    mutate(ar = ar_spectrum(pwelch_both$ff, phis, sd, delta)) %>%
    filter(m == "m=64") %>%
    ggplot() +
    geom_line(aes(x = ff, y = ar), linewidth = 1, alpha = 0.3) +
    geom_line(aes(x = ff, y = pwelch, colour = "pwelch")) +
    geom_line(
        data = dwelch_both %>% filter(m == "m=64"),
        mapping = aes(x = ff, y = dwelch, colour = "dwelch")
    ) +
    scale_x_continuous(
        "Frequency [Hz]",
        limits = c(0.25, 0.35),
        expand = c(0, 0)
    ) +
    scale_y_continuous(
        "Spectral density",
        limits = c(1e-2, 5e-1),
        expand = c(0, 0),
        trans = "log10"
    ) +
    scale_color_manual(
        values = c("pwelch" = "#002385", "dwelch" = "#8d3b00")
    ) +
    theme(legend.position = "none", axis.title = element_blank())

p1 +
    annotation_custom2(
        grob = ggplotGrob(p2a),
        data = data.frame(m = "m=16"),
        ymin = 0.5, ymax = 4, xmin = 0.21, xmax = 0.47
    ) +
    annotation_custom2(
        grob = ggplotGrob(p2b),
        data = data.frame(m = "m=64"),
        ymin = 0.5, ymax = 4, xmin = 0.21, xmax = 0.47
    ) +
    geom_rect(
        aes(xmin = 0.25, xmax = .35, ymin = 1e-2, ymax = 5e-1),
        colour = "black", linetype = "dashed", alpha = 0
    )

```